---
title: "ID Modeling Final Project"
author: "Ashley Tseng (at3346), Alicia Yang (ay2452), Cassandra Coulter (cc4347), and Christopher Armbruster (cwa2118)"
date: "Due: May 6, 2020"
output: html_document
---

## For our EPID P8477: Epi Modeling for Infectious Diseases final project, we will be re-creating a SEIR/SIRS model for the 2010 cholera otbreak in Haiti with a vaccination intervention, then applying the model to another historical outbreak.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


## SEIR Model
```{r seir}
# Based on SEIR model from P8477 lab #3
SEIR = function(t,state,parameters){
  with(as.list(c(state,parameters)),{
    
    dS = -beta*S*I/N;
    dE = beta*S*I/N-alpha*E;
    dI = alpha*E-gamma*I;
    
    dcumInci = beta*S*I/N;
    
    list(c(dS, dE, dI, dcumInci))
  })
}

# Specify initial conditions/parameters
times_SEIR = 0:100
N = 1e5; 
E0 = 10; 
I0 = 0; 
S0 = N-E0-I0;
state_SEIR = c(S = S0, E = E0, I = I0, cumInci = 0);
parameters_SEIR = c(beta = 0.6, alpha = 1/1.5, gamma = 1/3); 
sim_SEIR = ode(y = state_SEIR, times = times_SEIR, func = SEIR, parms = parameters_SEIR)
```



## SIRS Model
### Is a SIRS model the same as a SIR model? Do we need to incorporate a second "S" equation in our model for SIRS?
### SIRS is a little different from SIR--I think it takes into account waning immunity. For our overall model, we may want to include another "S" but depending on the scenario we choose and the length of the outbreak, I'm not sure if we'll need to take waning immunity into account for our scenario.
```{r sirs}
# Based on SIR model from P8477 lab #2
SIR = function(t,state,parameters){
  with(as.list(c(state,parameters)),{
    
    dS = -beta*S*I/N;
    dI = beta*S*I/N-gamma*I;
    dR = gamma*I;
    
    list(c(dS,dI,dR))
  })
}

# Specify initial conditions/parameters
times_SIR = seq(0, 365, by = 1);  # run for a year
N = 1e5; # population size
I0 = 10; # initial No. of Infectious people
S0 = N - I0;  # initial No. of Susceptible people
R0 = 0; # initial No. of Recovered people
state_SIR = c(S = S0, I = I0, R = R0)
parameters_SIR = c(beta = 0.5, gamma = 0.3)
sim_SIR = ode(y = state_SIR, times = times_SIR, func = SIR, parms = parameters_SIR)
```



## SIAR model w/ W & V

### The following notations for the compartments were used by Kirpich et al. (2017) (see: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5406029/):
### S(t)—the size of the susceptible compartment at a given time t.
### A(t)—the size of the asymptomatic compartment at a given time t.
### I(t)—the size of the symptomatic compartment at a given time t.
### R(t)—the size of the recovered compartment at a given time t.
### V(t)—the size of the vaccinated compartment at a given time t.
### W(t)—bacteria concentration in water at a given time t (environmental compartment).
```{r siar}
# How do we simplify the intervention model equations from Kirpich et al. (2017) so that we can code them in R Studio?
#Current equations here are directly taken from Kirpich et al.

SIAR = function(t,state,parameters){
  with(as.list(c(state,parameters)),{
    
    dS = muh + N + murs*R + muvs*V - (muwsa + muwsi)*S*(W/(kappa+W)) - muhsa + muhsi*S*A + I -(musv +muh_)*S;
    dA = (muwsa*S + muwva*V)*(W/(kappa+W)) + (muhsa*S + muhva*V)*(A+I-(muar + muh_));
    dI = (muwsi*S + muwvi*V)*(W/(kappa+W)) + (muhsi*S +muhvi*V)*(A+I-(muir + muh_))*I;
    dR = muar*A + muir*I - (murs + muh_)*R;
    dV = musv*S - (muwva + muwvi)*V*(W/(kappa+W)) - (muhva + muhvi)*V*A+I-(muvs+muh_)*V;
    dW = g*(muaw*A + muiw*I + h*M*W-gammaw_*W);
    
    list(c(dS,dA,dI,dR,dV,dW))
  })
}
```

